---
title: "individual1"
author: "Oztok"
date: "3/3/2021"
output: html_document
---

```{r setup, include=FALSE}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
```

#LOAD
```{r}
dat <- readxl::read_xls('GSS.xls')
library(classdata)
library(ggplot2)
library(dplyr)
library(tidyverse)
```



##Rename
```{r}
dat <- dat %>% rename(year = `Gss year for this respondent`)
dat <- dat %>% rename(wrkslf = `R self-emp or works for somebody`)
dat <- dat %>% rename(wrkgovt = `Govt or private employee`)
dat <- dat %>% rename(marital = `Marital status`)
dat <- dat %>% rename(educ = `Highest year of school completed`)
dat <- dat %>% rename(sex = `Respondents sex`)
dat <- dat %>% rename(rincome = `Respondents income`)
dat <- dat %>% rename(partyid = `Political party affiliation`)
dat <- dat %>% rename(happy = `General happiness`)
```

<br>
<br>

<br>
<br>

##1
```{r}
dat1 <- dat[- grep("Don't know", dat$educ),]
dat2 <- dat1[- grep("No answer", dat1$educ),]
dat3 <- dat2[- grep("No answer", dat2$marital),]
dat3$educ <- as.numeric(as.character(dat3$educ))
dat3 %>%                                      
  group_by(marital) %>%                         
  summarise_at(vars(educ),              
               list(name = mean)) %>%
  arrange(desc(name))
```
<br>
<br>
## We found that people who were never married have the highest years of school experience
<br>
## We also found that widowed individuals have the lowest years of school experience.

<br>
<br>



##2
```{r}
better_educated <- dat3 %>% group_by(marital) %>% filter(educ > mean(educ, na.rm = TRUE))
str(better_educated)
```
<br>
<br>



##3
```{r}
dat4 <- dat3
happiness <- dat4[dat4$happy %in% c('Very happy' , 'Pretty happy'), ]

myTable = table(happiness$marital, happiness$happy)
myTable
myFrame <- as.data.frame(myTable) 


a <- ggplot(myFrame, aes(x = Var1, y = Freq, group = Var2)) +
  geom_bar(aes(fill = Var2), stat = "identity", position = "dodge") + 
  labs(fill = "Happiness status") + labs(x= "Marital Status", y="Count")
a
```
<br>
<br>
## I have used the happy and marital status variables
<br>
## As we can see the number of married people are generally tend to be  pretty happy compare to other marital status'
<br>
## The number of seperated people is the least happy group compared to other marital status'
<br>
<br>


##4
```{r}
happiness1 <- happiness[- grep("Don't know", happiness$partyid),]
happiness2 <- happiness1[- grep("No answer", happiness1$partyid),]
myTable2 = table(happiness2$marital, happiness2$happy, happiness2$partyid)
myTable2
myFrame2 <- as.data.frame(myTable2)
ggplot(myFrame2, aes(x=Var1, weight= Freq)) + geom_bar() + facet_grid(Var3 ~ Var2 )
```
<br>
I do not believe there is a strong relation between affiliation between marital status, happiness. It is distributed pretty evenly.
<br>
<br>

##5
```{r}
dat5 <- dat3
myTable3 = table(dat5$marital, dat5$educ)
myTable3

myFrame3 <- as.data.frame(myTable3) 

z <- ggplot(myFrame3, aes(x = Var2, y = Freq, group = Var1)) +
  geom_bar(aes(fill = Var1), stat = "identity", position = "dodge") + 
  labs(fill = "Happiness status") + labs(x= "Years of school", y="Count") + coord_flip()
z
```
<br>
<br>
As seen in the graph, the # people who had 12 years of education or more tend to be married compared to the other education levels.
<br>
<br>

##6
```{r}
sector <- happiness
sector_h <- sector[sector$wrkgovt %in% c('Government' , 'Private'), ]


myTable6 = table(sector_h$wrkgovt, sector_h$happy)
myTable6
myFrame6 <- as.data.frame(myTable6) 


a <- ggplot(myFrame6, aes(x = Var1, y = Freq, group = Var2)) +
  geom_bar(aes(fill = Var2), stat = "identity", position = "dodge") + 
  labs(fill = "Happiness status") + labs(x= "Government or Private employee", y="Count")
a

```
<br>
<br>
It is very obvious to see the # of people who work in private industry is greater than # of people who work in government.
<br>
People tend to be pretty happy instead of very happy in both categories
<br>
<br>



##7
```{r}
dat7 <- dat
wrk_self <- dat7[dat7$wrkslf %in% c('Self-employed' , 'Someone else'), ]
income_comp <- wrk_self[wrk_self$rincome %in% c('Lt $1000' , '$1000 to 2999',  '$3000 to 3999',  '$4000 to 4999', '$5000 to 5999', '$6000 to 6999', '$7000 to 7999', '$8000 to 9999', 
                                              '$10000 - 14999', '$15000 - 19999', '$20000 - 24999', '$25000 or more'), ]

myTable7 = table(income_comp$wrkslf, income_comp$rincome)
myTable7
myFrame7 <- as.data.frame(myTable7) 


o <- ggplot(myFrame7, aes(x = Var1, y = Freq, group = Var2)) +
  geom_bar(aes(fill = Var2), stat = "identity", position = "dodge") + 
  labs(fill = "Self-Employed or Work for Someone Else") + labs(x= "Income Leves", y="Count")
o

```
<br>
<br>
This one was very surprising for me so I had to check my code multiple times
<br>
People who work for someone else gets paid more in every single income bracket
<br>
<br>

